# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv", header = T)
head(activity)
```

```{r}
activity$date <- as.Date(activity$date)
activity$weekdays <- weekdays(activity$date)
activity$wfactor <- activity$weekdays == "Sunday" | activity$weekdays == "Saturday"
activity$wfactor[activity$wfactor == T] <- "Weekend"
activity$wfactor[activity$wfactor == F] <- "Weekday"

library(reshape2)
activityMelt <- melt(activity, c("date", "wfactor", "interval"), c("steps"), na.rm = T)

#Q1: number steps taken by day
stepsdata <- dcast(activityMelt, date ~ variable, sum)

#Q2-Q4: Average  number of steps (across days) by interval
avgdata <- dcast(activityMelt, interval + wfactor ~ variable, mean)

```

## What is mean total number of steps taken per day?

```{r}
hist(stepsdata$steps, main = "Frequency of number of steps taken each day", xlab = "Number of Steps", breaks = 30, col = "red")
```

The above distribution mean and median are:

```{r}
mean <- mean(stepsdata$steps)
median <- median(stepsdata$steps)
```

```{r}
mean
median
```

## What is the average daily activity pattern?

Time series plot of the 5 min interval and the average number of steps across days:

```{r}
plot(avgdata$interval, avgdata$steps, type = "l",
     xlab = "Interval",
     ylab = "Average Number of Steps",
     main = "Average Number of Steps across Days by Interval",
     col = "red")
```

The 5-min interval that contains, on average, the maximum number of steps is:

```{r}
max <- avgdata[which.max(avgdata$steps),"interval"]
max
```

## Imputing missing values

```{r}

copy <- activity

for(i in which(is.na(copy$steps))) {
      int <- copy[i, "interval"]
      copy[i, "steps"] <- mean(avgdata[avgdata$interval == int, "steps"])
      }

head(copy)

```

Now, let's calculate the sum of steps taken by day and generate a histogram using this "copy" dataset (which looks like the activity dataset used in Q1, but with the missing values filled in). Then, we can check the mean and median values.

```{r}
copyMelt <- melt(copy, c("date","interval"), "steps", na.rm = T)
stepscopy <- dcast(copyMelt, date ~ variable, sum)

hist(stepscopy$steps, main = "Frequency of number of steps taken each day", xlab = "Number of Steps", breaks = 30, col = "blue")

mean(stepscopy$steps)
median(stepscopy$steps)

```

Apparently, inputting averages do not change dramatically either the graph or the distribution mean and median (11000ish).

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(ggplot2)
g <- ggplot(avgdata, aes(interval, steps))
g + geom_line() + facet_grid(wfactor~.) + theme_bw()